Appendix

### 2025-03-28 manuscript_rmarkdown/manuscript_appendix.Rmd
Added table for comparing estimators on 10000 randomly sampled datasets.

@@

>>>>

These transformed data

@@

# Yeo-Johnson on clean and dirty data.
clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")

data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))

# Add rank based on residual error for central normality.
data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]

table_data <- data[
  ,
  list(
    "avg." = round(mean(method_rank), digits = 2L),
    "10%" = quantile(method_rank, 0.10),
    "50%" = quantile(method_rank, 0.50),
    "90%" = quantile(method_rank, 0.90)
  ),
  by = c("transformation", "outlier", "dataset")
]

data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")

```
```{=latex}
\begin{table}
\begin{center}
\caption{
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:comparison_methods_simulations}
\begin{tabular}{l | l r r r r r r}

\toprule
& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
transformation & outliers: & no & yes & no & yes & no & yes \\

none                                  & & 8.51          & 6.33          & 7.31          & 5.55          & 7.42          & 6.39 \\
conventional                          & & 3.72          & 3.78          & 3.83          & 4.74          & 6.47          & 5.47 \\
conventional (z-score norm.)          & & 4.72          & 2.89          & 6.37          & 3.43          & 5.42          & 3.50 \\
conventional (robust scaling)         & & 4.07          & 4.31          & 5.36          & 4.23          & 4.41          & 4.32 \\
Raymaekers-Rousseeuw                  & & 4.50          & 7.03          & 3.97          & 8.12          & 6.25          & 5.57 \\
Raymaekers-Rousseeuw (z-score norm.)  & & 5.26          & 6.36          & 6.22          & 6.42          & 5.26          & 6.92 \\
Raymaekers-Rousseeuw (robust scaling) & & 4.71          & 7.84          & 5.31          & 7.06          & 4.35          & 7.58 \\
invariant                             & & \textbf{3.15} & 3.72          & \textbf{3.27} & 3.14          & \textbf{2.69} & 3.05 \\
robust invariant                      & & 6.37          & \textbf{2.73} & 3.36          & \textbf{2.30} & 2.72          & \textbf{2.20} \\

\bottomrule
\end{tabular}
\end{center}
\end{table}
```

# Yeo-Johnson on clean and dirty data.
clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean", method = "box_cox")
dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty", method = "box_cox")
shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted", method = "box_cox")

data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))

# Add rank based on residual error for central normality.
data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]

table_data <- data[
  ,
  list(
    "avg." = round(mean(method_rank), digits = 2L),
    "10%" = quantile(method_rank, 0.10),
    "50%" = quantile(method_rank, 0.50),
    "90%" = quantile(method_rank, 0.90)
  ),
  by = c("transformation", "outlier", "dataset")
]

data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")

### 2025-03-28 manuscript_rmarkdown/manuscript_appendix.tex
Added table for comparing estimators on 10000 randomly sampled datasets.

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-03-27}

@@

@@

>>>>

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-03-28}

@@

Robust transformations are hypothesised to have a cost in efficiency for
data without outliers, i.e.~clean data. Here we randomly draw data from
a standard normal distribution \(\mathcal{N}(0,1)\). Subsequently, we
perform an inverse transformation
\(\left(\phi_{\text{YJ}}^{\lambda, 0, 1}\right)^{-1}\), with
\(\lambda \in \left(0, 2\right)\).

These transformed data

@@

\begin{table}
\begin{center}
\caption{
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
\label{tab:comparison_methods_simulations}
\begin{tabular}{l | r r r r r r}

\toprule
& \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
outliers & no & yes & no & yes & no & yes \\

none                                  & 8.51          & 6.33          & 7.31          & 5.55          & 7.42          & 6.39 \\
conventional                          & 3.72          & 3.78          & 3.83          & 4.74          & 6.47          & 5.47 \\
conventional (z-score norm.)          & 4.72          & 2.89          & 6.37          & 3.43          & 5.42          & 3.50 \\
conventional (robust scaling)         & 4.07          & 4.31          & 5.36          & 4.23          & 4.41          & 4.32 \\
Raymaekers-Rousseeuw                  & 4.50          & 7.03          & 3.97          & 8.12          & 6.25          & 5.57 \\
Raymaekers-Rousseeuw (z-score norm.)  & 5.26          & 6.36          & 6.22          & 6.42          & 5.26          & 6.92 \\
Raymaekers-Rousseeuw (robust scaling) & 4.71          & 7.84          & 5.31          & 7.06          & 4.35          & 7.58 \\
invariant                             & \textbf{3.15} & 3.72          & \textbf{3.27} & 3.14          & \textbf{2.69} & 3.05 \\
robust invariant                      & 6.37          & \textbf{2.73} & 3.36          & \textbf{2.30} & 2.72          & \textbf{2.20} \\

\bottomrule
\end{tabular}
\end{center}
\end{table}
